---
title: "Connecting Political Views and Health Metrics"
author: "Riyadh Baksh (rhb2152), Anika Mitchell (am5088), Jeong Yun Yang (jy3306)"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,message=FALSE,warning=FALSE)
```

## Motivation

The goal of this project is emulate an *ecological* study in order to understand and visualize the relationship between political trends and outcomes on county-level public health in the United States. Specifically, we are interested in the potential correlations between a variety of health metrics and presidential election results in both 2012 and 2016. By understanding the relationship between these data, this project aims to understand sociopolitical and socioeconomic implications for health and wellbeing at a county level, and offer insights into patterns that may offer recommendations for broader application.

## Related Work

The inspiration for this project comes from a growing  body of scientific literature aimed at understanding how health transforms as the sociopolitical climate shifts.

For example, Salas-Guerra’s 2021 multivariate correlational analysis of four U.S states (California, New York, Texas, and Florida) investigated political affiliation and viral contagion/mortality from COVID-19. Findings indicated poverty and lack of insurance as a main risk factor for the COVID contagion in Republican states (Salas-Guerra, 2021).

Other studies such as Bor’s 2017 study “*Diverging Life Expectancies and Voting Patterns in the 2016 US Presidential Election*” examined voting patterns and life expectancy at the county level. Main findings included voting patterns undermining the overall health of the voter; that is, a pattern was detected of residents who voted for Trump or didn’t vote at all in 2016 and were most likely to be residents of counties that experienced not only economic insecurity but decreased public health indicated by higher mortality rates (Bor 2017).

Curtis et al.’s 2021 study “*Life expectancy and voting patterns in the 2020 U.S presidential election*” came to similar conclusions leveraging county-level life expectancy data and voting data scraped from news outlets. Specifically, they found that despite large shifts towards Democratic candidates, communities that were most impacted by lower life expectancy and lower health tended to vote Republican (Curtis et al., 2021).

The results of these studies indicate a key point of interest related to what communities choose to vote for compared to what those implications on their health are. Our project aims to add to this body of data analysis with a specific focus on health indicators like premature death, preventable hospital stays, number of physicians geographically, and flu vaccinations.

## Initial Questions

Our main research question is: how do national and county level bipartisan voting patterns impact key health indicators (premature death, preventable hospital stays, number of physicians, and vaccination status) in 2012 and 2016. Our original research question was interested in assessing this relationship from 2000-2020, however we chose to constrain our data to using only 2012 and 2016 data after conducting early data wrangling and exploratory analyses that identified this set of time as having the most robust dataset available to appropriately assess our relationship of interest. Additionally, this date range (2012-2016) is of particular public health interest, as the decades long trend of increased life expectancy between 1959 and 2016 started to noticeably slow down after 2014 (Woolf and Schoomaker, 2019).

## Data

For information on health outcomes by county and by year, we used information from this source: https://www.countyhealthrankings.org/health-data/methodology-and-sources/data-documentation

For data on election results by county and by year, we consulted the MIT election datasets: https://electionlab.mit.edu/data

```{r message=FALSE}
library(tidyverse)
```

We started by cleaning the `election` dataset, which contains information such as county name, state, year, party, and total votes. The mutate function was used to calculate the percentage of votes. The common variable `year_county` was used to uniquely match the data for a given year and county FIPS code.

```{r, message=FALSE}
election =
  read_csv("data/election.csv") |>
  filter(party=="DEMOCRAT" | party=="REPUBLICAN") |>
  group_by(county_fips,county_name,state_po,year,party,totalvotes) |>
  summarize(
    votes = sum(candidatevotes)
  ) |>
  mutate(
    percent = votes/totalvotes,
    county_fips = sprintf("%05d",county_fips),
    year_county = paste(year,county_fips,sep="-")
  ) |>
  select(-totalvotes,-votes) |>
  pivot_wider(
    names_from = "party",
    values_from = "percent"
  ) |>
  janitor::clean_names()
```

We identified four measures that are the most robust in the datasets: `Premature death`, `PReventable hospital stays`, `Primary care physicians`, and `flu vaccinations`, and filtered for these four measurements. Since these measurements have different units and scales, we standardize each variable to improve visualization and statistical analysis. This means that for each health metric, a value of zero means the average, and each unit represents one standard deviation away. For example, a `premature_death` value of 1 means that county has a premature death rate that is one standard deviation higher than the average county.  

```{r, message=FALSE}
health =
  read_csv("data/health.csv") |>
  mutate(
    year = substr(yearspan,start=1,stop=4),
    county_fips = paste(statecode,countycode,sep=""),
    year_county = paste(year,county_fips,sep="-")
  ) |>
  filter(
    measurename == "Premature death" |
    measurename == "Preventable hospital stays" |
    measurename == "Primary care physicians" |
    measurename == "Flu vaccinations"
  ) |>
  select(year_county,year,county_fips,county,state,measurename,rawvalue) |>
  pivot_wider(
    names_from = "measurename",
    values_from = "rawvalue"
  ) |>
  janitor::clean_names() |>
  mutate(
    premature_death = scale(premature_death),
    primary_care_physicians = scale(primary_care_physicians),
    preventable_hospital_stays = scale(preventable_hospital_stays),
    flu_vaccinations = scale(flu_vaccinations)
  )
```

Here we join the `election` and `health` datasets by a common variable `year_county`. Through our exploratory analysis, we found that years 2012 and 2016 had the least amount of missing data, hence we filter for data for these years. Then we remove the outliers for each measurements.

```{r}
health_election =
  left_join(election,health,by="year_county") |>
  rename(
    county_fips=county_fips.x,
    year=year.x) |>
  select(county_fips,county_name,state_po,year,democrat,republican,
         premature_death,
         physicians=primary_care_physicians,
         preventable_hospital_stays,
         flu_vaccinations) |>
  filter(year==2012 | year==2016)

health_election_with_na = health_election

health_election = drop_na(health_election)

# remove outliers in premature death
quartiles = quantile(health_election$premature_death, probs=c(.25,.75))
IQR = IQR(health_election$premature_death)

lower = quartiles[1] - 1.5*IQR
upper = quartiles[2] + 1.5*IQR

health_election =
  health_election |>
  filter(
    premature_death > lower &
    premature_death < upper
  )

# remove outliers in primary care physicians
quartiles = quantile(health_election$physicians, probs=c(.25,.75))
IQR = IQR(health_election$physicians)

lower = quartiles[1] - 1.5*IQR
upper = quartiles[2] + 1.5*IQR

health_election =
  health_election |>
  filter(
    physicians > lower &
    physicians < upper
  )

# remove outliers in preventable hospital stays
quartiles = quantile(health_election$preventable_hospital_stays, probs=c(.25,.75))
IQR = IQR(health_election$preventable_hospital_stays)

lower = quartiles[1] - 1.5*IQR
upper = quartiles[2] + 1.5*IQR

health_election =
  health_election |>
  filter(
    preventable_hospital_stays > lower &
    preventable_hospital_stays < upper
  )

# remove outliers in flu vaccinations
quartiles = quantile(health_election$flu_vaccinations, probs=c(.25,.75))
IQR = IQR(health_election$flu_vaccinations)

lower = quartiles[1] - 1.5*IQR
upper = quartiles[2] + 1.5*IQR

health_election =
  health_election |>
  filter(
    flu_vaccinations > lower &
    flu_vaccinations < upper
  )
```

## Exploratory Analysis

We first started by exploring each dataset individually. For the `election` data, a plot was made comparing the county vote share for `democrat` and for `republican`. It was observed that in many counties, the sum of the `democrat` and `republican` percents do not add to 1, which indicates the presence of third-parties. That being said, the graph below shows there's a limit where no points are above the line $y=x$. This makes sense because the vote share in a specific county cannot exceed 100%.

```{r}
election |>
  drop_na() |>
  ggplot(aes(x=democrat,y=republican)) + geom_point()
```

Next, we considered how the four health variables under investigation are related to each other. This was important to make sure they are all independent from each other. In general, there is largely random distribution of points, so there is no clear pattern. However, the relationship between `premature_death` and `preventable_hospital_stays` does suggest a positive correlation. Counties with higher premature death rates also seem to have more preventable hospital stays, which makes sense because a premature death means by definition that it was a preventable death. There also seems to be a negative correlation between `premature_death` and `flu_vaccinations`. With greater flu vaccinations, we observe fewer premature deaths.

```{r}
library(patchwork)

p1 =
  health_election |>
  ggplot(aes(x=premature_death,y=physicians)) +
  geom_point(size=0.5)

p2 =
  health_election |>
  ggplot(aes(x=premature_death,y=preventable_hospital_stays)) +
  geom_point(size=0.5)

p3 =
  health_election |>
  ggplot(aes(x=premature_death,y=flu_vaccinations)) +
  geom_point(size=0.5)

p4 =
  health_election |>
  ggplot(aes(x=physicians,y=preventable_hospital_stays)) +
  geom_point(size=0.5)

p5 =
  health_election |>
  ggplot(aes(x=physicians,y=flu_vaccinations)) +
  geom_point(size=0.5)

p6 =
  health_election |>
  ggplot(aes(x=preventable_hospital_stays,y=flu_vaccinations)) +
  geom_point(size=0.5)

(p1 + p2 + p3) / (p4 + p5 + p6)


```

Now, we began to look at the correlation between election results and these four health metrics. For simplicity, the outcome or response variable was chosen to be the `republican` vote share. However, as we demonstrated above, this is negatively correlated with the `democrat` vote share.

There does seem to be a lot of random scatter and a clear pattern is hard to justify. However, some interesting observations are the positive correlations between `physicians` and `republican` vote share. There also seems to be a negative correlation between `flu_vaccinations` and `republican` vote share.

```{r}
a1 =
  health_election |>
  ggplot(aes(x=premature_death,y=republican)) + geom_point(size=0.5)

a2 =
  health_election |>
  ggplot(aes(x=physicians,y=republican)) + geom_point(size=0.5)

a3 =
  health_election |>
  ggplot(aes(x=preventable_hospital_stays,y=republican)) + geom_point(size=0.5)

a4 =
  health_election |>
  ggplot(aes(x=flu_vaccinations,y=republican)) + geom_point(size=0.5)

(a1 + a2)/(a3 + a4)
```

### Geographic Plots

We also decided to include maps as part of the exploratory analysis. The plots below show the nationwide county-level vote share. The darker blue represents higher `democrat` vote share. The darker red represents higher `republican` vote share. As expected, these are inversely related - that is, a darker blue area would also be lighter red. To improve comparability, the scales are fixed so that the same shade of blue, for example, means the same thing between the two years. We noticed that when going from 2012 to 2016, the map became a lot lighter blue, reflecting the decrease in the democrat popular vote from 2012 to 2016. Note: Alaska was removed due to a different way of assigning counties, and thus Hawaii was also removed so that we only considered the contiguous US.


```{r, message=FALSE}
#devtools::install_github("UrbanInstitute/urbnmapr")
#remotes::install_github("UrbanInstitute/urbnthemes",build_vignettes = TRUE)

library(urbnmapr)
library(urbnthemes)
```

```{r}
library(urbnmapr)

# Plots for 2012
county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_name!="Alaska" & state_name!="Hawaii")

dem_2012 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = democrat)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#f5fbff",
                      high="#4169e1",
                      limits=c(0.05,0.95)) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group),
               fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "democrat") +
  theme_urbn_map()

rep_2012 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = republican)) +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#fdf5f5",
                      high="#dc2323",
                      limits=c(0.05,0.95)) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group),
               fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "republican") +
  theme_urbn_map()

# Plots for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_name!="Alaska" & state_name!="Hawaii")

dem_2016 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = democrat)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#f5fbff",
                      high="#4169e1",
                      limits=c(0.05,0.95)) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group),
               fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "democrat") +
  theme_urbn_map()

rep_2016 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = republican)) +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#fdf5f5",
                      high="#dc2323",
                      limits=c(0.05,0.95)) +
  geom_polygon(data = states, mapping = aes(long, lat, group = group),
               fill = NA, color = "#ffffff") +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "republican") +
  theme_urbn_map()

(dem_2012+dem_2016)/(rep_2012+rep_2016)
```

While the intention was to look at nationwide correlations, we realized that there are many factors that influence voting across an entire country. It is also more difficult to identify trends in large heterogeneous populations. Therefore, we decided to also zoom in on New York state. We also observed that from 2012 to 2016, the state became lighter blue and darker red.

```{r,warning=FALSE}
# Plots for 2012

county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

dem_2012 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = democrat)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#f5fbff",
                      high="#4169e1",
                      limits=c(0.2,0.9)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "democrat") +
  theme_urbn_map()

rep_2012 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = republican)) +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#fdf5f5",
                      high="#dc2323",
                      limits=c(0.05,0.95)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "republican") +
  theme_urbn_map()

# Plots for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

dem_2016 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = democrat)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#f5fbff",
                      high="#4169e1",
                      limits=c(0.2,0.9)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "democrat") +
  theme_urbn_map()

rep_2016 =
county_data |>
  ggplot(aes(long, lat, group = group, fill = republican)) +
  geom_polygon(color = NA) +
  scale_fill_gradient(labels = scales::percent,
                      guide = "colourbar",
                      low="#fdf5f5",
                      high="#dc2323",
                      limits=c(0.05,0.95)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
        legend.key.width = unit(.5, "in")) +
  labs(fill = "republican") +
  theme_urbn_map()

(dem_2012+dem_2016)/(rep_2012+rep_2016)
```

Next, we considered the geographic distributions of the health metrics. It would be difficult to see trends across the entire US, so instead, we are only looking at New York state.

There does not seem to be a clear pattern for `premature_death`, as some counties became lighter whereas others became darker.

```{r}
# PREMATURE DEATH

# Plot for 2012
county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2012 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = premature_death)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-2,0.5)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "premature_death") +
  theme_urbn_map()

# Plot for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2016 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = premature_death)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-2,0.5)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "premature_death") +
  theme_urbn_map()

plot_2012 + plot_2016
```

The `physicians` rate in each county appears to stay roughly constant when going from 2012 to 2016.

```{r}
# PHYSICIANS

# Plot for 2012
county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2012 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = physicians)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1,1.2)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "physicians") +
  theme_urbn_map()

# Plot for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2016 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = physicians)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1,1.2)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "physicians") +
  theme_urbn_map()

plot_2012 + plot_2016
```

At first glance, it seems that there were fewer `preventable_hospital_stays` in 2016 compared to 2012.

```{r}
# PREVENTABLE HOSPITAL STAYS

# Plot for 2012
county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2012 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = preventable_hospital_stays)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1.2,1.8)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "preventable_hospital_stays") +
  theme_urbn_map()

# Plot for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2016 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = preventable_hospital_stays)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1.2,1.8)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "preventable_hospital_stays") +
  theme_urbn_map()

plot_2012 + plot_2016
```

Further, `flu_vaccinations` seems to have stayed roughly constant from 2012 to 2016.

```{r}
# FLU VACCINATIONS

# Plot for 2012
county_data =
  health_election_with_na |>
  filter(year==2012) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2012 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = flu_vaccinations)) +
  ggtitle("2012") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1,1.4)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "flu_vaccinations") +
  theme_urbn_map()

# Plot for 2016
county_data =
  health_election_with_na |>
  filter(year==2016) |>
  right_join(counties,by="county_fips") |>
  filter(state_po=="NY")

plot_2016 =
  county_data |>
  ggplot(aes(long, lat, group = group, fill = flu_vaccinations)) +
  ggtitle("2016") +
  geom_polygon(color = NA) +
  scale_fill_gradient(guide = "colourbar",
                      low="#cce7c9",
                      high="#276221",
                      limits=c(-1,1.4)) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  theme(legend.title = element_text(),
  legend.key.width = unit(.5, "in")) +
  labs(fill = "flu_vaccinations") +
  theme_urbn_map()

plot_2012 + plot_2016
```

## Additional Analysis

Next, we conducted a series of multiple linear regressions to quantify the effect of the health metrics on the `republican` vote share. There were a total of six regressions: the US for both years, the US for 2012, the US for 2016, NY state for both years, NY state for 2012, and NY state for 2016. The estimates for the coefficients and the corresponding R2 values are summarized in the tables below. Note that while not all estimates were significant, most were at the 5% level. These numbers will be further discussed in the next section.

```{r}
overall = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = health_election)

overall_2012 = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = filter(health_election,year==2012))

overall_2016 = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = filter(health_election,year==2016))

NY = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = filter(health_election,state_po=="NY"))

NY_2012 = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = filter(health_election,state_po=="NY",year==2012))

NY_2016 = lm(republican ~ premature_death + physicians + preventable_hospital_stays + flu_vaccinations, data = filter(health_election,state_po=="NY",year==2016))

reg = tibble(
  term=broom::tidy(overall)$term,
  US_Overall=overall$coefficients,
  US_2012=overall_2012$coefficients,
  US_2016=overall_2016$coefficients,
  NY_Overall=NY$coefficients,
  NY_2012=NY_2012$coefficients,
  NY_2016=NY_2016$coefficients
  )
knitr::kable(reg, digits=3,caption="Regression Coefficients on Percent Republican")

r2 = tibble(
  US_Overall=broom::glance(overall)$r.squared,
  US_2012=broom::glance(overall_2012)$r.squared,
  US_2016=broom::glance(overall_2016)$r.squared,
  NY_Overall=broom::glance(NY)$r.squared,
  NY_2012=broom::glance(NY_2012)$r.squared,
  NY_2016=broom::glance(NY_2016)$r.squared
  )
knitr::kable(r2, digits=3,caption="R2 Values")
```


## Discussion



